|

### 2023-01-17 paper_comparison/paper_comparison/paper_comparison.Rmd
update main sections

```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example fitted vs residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "Breusch-Pagan", 
                             "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$. Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup. Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$. If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.

@@

The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two regressors. Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin. This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals. (expand?)

@@

Data used in the $m - 1$ null plots needs to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

In hypothesis testing, $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true. Within the context of visual inference, by involving $k$ independent observers, $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.

Let $X_j = \{0,1\}$ be a binomial random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].

@@

The power of a model misspecification test specifies the chance of a misspecified regression model being detected given a certain alternative hypothesis. It is an important indicator when one concerns if the regression model is misspecified. Though, people might usually be more interested in knowing how much the residuals depart from the model assumptions and if the departure is of practical significance.

As discussed in @majumder_validation_2013, individual's skill may affect the number of observers who identify the data plot from the lineup. Thus, the power of a visual test could depend on the subject-specific abilities. Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, in the multiple selections scenario, having this probability is insufficient to determine the power of a visual test as it doesn't provide information about the number of selections made by the subject for p-value calculation.

Instead, we estimate the probability of a lineup being rejected directly by assuming the individual skill has negligible impact on the variation of the power of a lineup. The assumption is made to simplify the model structure, thus mitigating the need for expensive large-scale experiments to estimate the complex covariance matrix. The model is a logistic regression with the natural logarithm of the effect as the only regressor formulated as

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(E)),
\end{equation}
```
\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

@@

```{=tex}
\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}
```
\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. This includes [expansion].

@@

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{x}, 2)),
\end{align}

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), \\
\end{align}
```

@@

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level. Uniform distribution is the focus of this section because visual patterns are more likely to be revealed. Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable power curves. Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}. At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$. Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.

@@

For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that REST test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable visual features can be spotted from the example residual plots.

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
             group_by(log_effect_size, reject) %>% 
             summarise(num_response = n()),
           aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3)  +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_size_manual(values = c(1, 2, 3)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Effect size (log e)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

@@

Depending on the natural logarithm of the effect, lineups are classified into three categories: easy (> 3.5), moderate (1.5 - 3.5) and difficult (< 1.5). Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests at different difficulty levels.

### Easy lineup

For easy lineups produced by the non-linearity model, 65% are rejected by both the conventional and visual tests, while 9% are accepted by both tests. Thus, 74% of the time two tests will agree with each other. There are 16% of lineups rejected by conventional tests only.

Interestingly, in spite of the greater power of conventional tests, 10% of the lineups are rejected by visual tests only. Recall that the RESET test used in our analysis includes powers of fitted values up to four. However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively. In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected. We also can observe from the example plot displayed in Figure \ref{fig:easy-poly-example} that panel three is distinctly different from others, exhibiting a "M" shape. This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.

```{r easy-poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III. It is rejected by the visual test but not by the RESET test. The data plot at panel three exhibits strong non-linear pattern."}
VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

For easy lineups produced by the heteroskedasticity model, 62% are rejected by both the conventional test and the visual test, and 7% are accepted by both, result in 69% agreement rate. Lineups only rejected by conventional test are 30%. 1% of lineups are accepted by the conventional test but rejected by the visual test.

Among those 1%, which are two lineups, one shows detectable heteroskedasticity pattern as displayed in Figure \ref{fig:easy-heter-example1}. Another contain data plot with an noticeable outlier, but the "Right-triangle" shape is not very strong as shown in Figure \ref{fig:easy-heter-example2}. The $p$-values of Breusch-Pagan test for these two lineups are $0.055$ and $0.135$. Compared to the visual test, its $p$-values are $0.026$ and $0.012$. The Breusch-Pagan test should reject the one with obvious heteroskedasticity signal, but it yields a $p$-value slightly above 5%.

```{r easy-heter-example1, fig.height = 5, fig.cap = 'Lineup heter-331 with $log_e(E) = 4.02$ produced by the heteroskedasticity model used in experiment II and III. It is rejected by the visual test but not by the  Breusch-Pagan test. The data plot at panel 17 contains a "Butterfly" shape, indicating the presence of heteroskedasticity.'}
VI_MODEL$plot_lineup(vi_lineup$heter_331$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

```{r easy-heter-example2, fig.height = 5, fig.cap = "Lineup heter-536 with $log_e(E) = 4.13$ produced by the heteroskedasticity model used in experiment II. It is rejected by the visual test but not by the  Breusch-Pagan test. The data plot at panel 18 exhibits a vague heteroskedasticity pattern, but it contains a distinct outlier which makes it attractive to subjects."}
VI_MODEL$plot_lineup(vi_lineup$heter_536$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

### Moderate lineup

>>>>

```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "Breusch-Pagan", 
                             "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the data plot, and the remaining $m - 1$ null plots have the identical graphical procedure except the data has been replaced with data consistent with $H_0$. Then, an observer who have not seen the data plot will be asked to point out the most different plot from the lineup. Under $H_0$, it is expected that the data plot would have no distinguishable difference from the null plots, and the probability that the observer correctly picks the data plot is $1/m$. If one rejects $H_0$ as the observer correctly picks the data plot, then the Type I error of this test is $1/m$.

@@

The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two regressors. Their results suggest that visual tests are capable of testing the significance of a single regressor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test. (contradict to our claim? expand?) In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin. This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals. (expand?)

@@

Data used in the $m - 1$ null plots needs to be simulated. In regression diagnostics, sampling data consistent with $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually a composite hypothesis controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to a so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

In hypothesis testing, a $p$-value is defined as the probability of observing test results as least as extreme as the observed result given $H_0$ is true. Within the context of visual inference, by involving $k$ independent observers, the $p$-value can be interpreted as the probability of having as many or more subjects detect the data plot than the observed result.

Let $X_j = \{0,1\}$ be a Bernoulli random variable denoting whether subject $j$ correctly detecting the data plot, and $X = \sum_{j=1}^{K}X_j$ be the number of observers correctly picking the data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the data plot [@majumder_validation_2013].

@@

The power of a model misspecification test is the probability that the null hypothesis is rejected given the regression model is misspecified. It is an important indicator when one is concerned about whether model assumptions have been violated. Although in practice, one might be more interested in knowing how much the residuals deviate from the model assumptions, and whether this deviation is of practical significance.

As discussed in @majumder_validation_2013, the power of a visual test may depend on the ability of the particular subject, as the skill of the individual may affect the number of observers who identify the data plot from the lineup. Previously, it is addressed by modelling the probability of a subject $j$ correctly picking the data plot from a lineup $l$ using a mixed-effect logistic regression, with subjects treated as random effects [@majumder_validation_2013]. However, in the multiple selections scenario, having this probability is not sufficient to determine the power of a visual test because it does not provide information about the number of selections made by the subject for the calculation of the p-value.

Instead, we directly estimate the probability of a lineup being rejected by assuming that individual skill has negligible effect on the variation of the power. The assumption is to simplify the model structure, thereby obviate costly large-scale experiments to estimate complex covariance matrices. The model is a logistic regression with the natural logarithm of the effect as the only regressor formulated as

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log(E)),
\end{equation}
```
\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

@@

\small

\begin{gather}  \label{eq:effect-size}
E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V})|}{|diag(\boldsymbol{R}_a\sigma^2)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V})^{-1}diag(\boldsymbol{R}_a\sigma^2)) + \boldsymbol{\mu}_z^{T}(\boldsymbol{R}_a\boldsymbol{V})^{-1}\boldsymbol{\mu}_z\right),\\
\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{X}_a(\boldsymbol{X}_a^T\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a^T,
\end{gather}

\normalsize

\noindent where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the model and $\boldsymbol{\beta}_z$ be the corresponding coefficients.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. These include the distribution of the regressor and the type of the simulated model.

@@

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align}

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}), 
\end{align}
```

@@

In the following analysis, lineups with $X_{raw} \sim U(-1,1)$ will be the focus of next two sections, because visual patterns are more likely to be revealed with uniform distribution. Besides, six more evaluations are collected for the uniform distribution in experiment III, which will produce more reliable and stable results. Analysis of other distributions can be found in Section \ref{effect-of-the-distribution-of-the-regressor}.

@@

Figure \ref{fig:polypower} shows the estimated power of visual test on lineups produced by the non-linearity model with $X_{raw} \sim U(-1,1)$, against the natural logarithm of the effect $log_e(E)$, with a 5% significance level. At the bottom of the figure \ref{fig:polypower}, there are a sequence of example residual plots with increasing levels of $log_e(E)$. Readers can evaluate them from left to right and determine at which level the departure from a good residual plot becomes detectable.

@@

For the non-linearity model, the power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable visual features can be spotted from the example residual plots.

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
             group_by(log_effect_size, reject) %>% 
             summarise(num_response = n()),
           aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3)  +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                filter(x_dist == "uniform") %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_size_manual(values = c(1, 2, 3)) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  xlab("Effect size (log E)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

```{r}
p <- vi_survey %>%
  filter(x_dist == "uniform") %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(reject = as.numeric(p_value <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_point(data = ~.x %>% 
               group_by(log_effect_size, reject) %>% 
               summarise(num_response = n()),
             aes(log_effect_size, reject, size = factor(num_response)), alpha = 0.3) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                filter(x_dist == "uniform") %>%
                select(-(RESET5_p_value:RESET10_p_value)) %>%
                select(-RESET3_p_value, -F_p_value) %>%
                rename(RESET_p_value = RESET4_p_value) %>%
                pivot_longer(RESET_p_value:SW_p_value) %>%
                mutate(name = gsub("_p_value", " test", name)), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject, times = 500), 
              aes(log_effect_size, reject, group = boot_id, col = "Visual test"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(0, 5, 0.5), minor_breaks = NULL) +
  scale_size_manual(values = c(1, 2, 3)) +
  xlab("Effect size (log E)") +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

```{r p-value-comparison, fig.width=10, fig.height=5, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model. Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}

p_value_cmp_dat <- vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  filter(x_dist == "uniform") %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(type = factor(type, levels = c("non-linearity", "heteroskedasticity"))) %>%
  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not"),
         reject = ifelse(p_value < 0.05, "Reject", "Not")) %>%
  mutate(conv_reject = factor(conv_reject, levels = c("Reject", "Not")),
         reject = factor(reject, levels = c("Reject", "Not")))

p_value_cmp_dat %>%
  ggplot() +
  ggmosaic::geom_mosaic(aes(x = ggmosaic::product(reject, conv_reject), fill = reject)) +
  facet_grid(~type) +
  ylab("Visual tests") +
  xlab("Conventional tests") +
  labs(fill = "Conventional tests") +
  scale_fill_brewer("", palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "none") +
  coord_fixed()
```

(remove lineup id from caption)

@@

Figure \ref{fig:p-value-comparison} provides a mosaic plot showing the rejection rate of visual tests and conventional tests for both the non-linearity model and the heteroskedasticity model.

For lineups produced by the non-linearity model, 25% are rejected by both the conventional and visual tests, while 42% are accepted by both tests. Thus, 69% of the time two tests will agree with each other. There are 26% of lineups rejected by conventional tests only.

Interestingly, in spite of the greater power of conventional tests, 7% of the lineups are rejected by visual tests only. Recall that the RESET test used in our analysis includes powers of fitted values up to four. However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials contain power of the regressor up to six and 18 respectively. In fact, by including the power of fitted values up to six in the RESET test, all these cases will be rejected. We also can observe from the example plot displayed in Figure \ref{fig:poly-example} that panel three is distinctly different from others, exhibiting a "M" shape. This clearly suggests users of the RESET test and any other residual-based conventional tests that require the specification of variables of interest to check the residual plot before conducting the test, such that the correct choice of variables can be made.

```{r poly-example, fig.height = 5, fig.cap = "Lineup poly-270 with $log_e(E) = 3.89$ produced by the non-linearity model used in experiment I and III. It is rejected by the visual test but not by the RESET test. The data plot at panel three exhibits strong non-linear pattern."}
VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

```{r}
poly_conv_sim %>%
  filter(x_dist == "uniform", shape %in% c(3, 4)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  select(shape, log_effect_size, (RESET3_p_value:RESET9_p_value)) %>%
  pivot_longer(RESET3_p_value:RESET9_p_value) %>%
  mutate(name = gsub("_p_value", " ", name)) %>%
  mutate(name = gsub("RESET", "", name)) %>%
  mutate(name = gsub("test", "", name)) %>%
  mutate(reject = as.numeric(value < 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = name), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab("Effect size (log E)") +
  ylab("Power") +
  scale_color_manual(values = rev(rcartocolor::carto_pal(7, "RedOr"))) +
  theme_light() +
  labs(col = "RESET test order") +
  facet_wrap(~shape)
```

```{r}
heter_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(BP_p_value < 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab("Effect size (log E)") +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
  theme_light() +
  labs(col = "Heteroskedasticity shape") +
  ggtitle("Conventional test")
```

```{r}
vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "heteroskedasticity") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(a = c("left-triangle", "butterfly", "right-triangle")[a + 2]) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = factor(a)), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab("Effect size (log E)") +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(3, "Vivid")) +
  theme_light() +
  labs(col = "Heteroskedasticity shape") +
  ggtitle("Visual test")
```
